---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Yiwei Tang"
date: "2/10/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

There are three Denny's locations in Alaska.

```{r filter_dennys_AK}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are two La Quinta locations in Alaska.

```{r filter_lq_AK}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

### Exercise 2

We need to calculate six locations between the two La Quintas and the three Denny's.

```{r calc_pair}
pair <- nrow(lq_ak) * nrow(dn_ak)
```

### Exercise 3

```{r join_sets}
dn_lq_ak <- full_join(dn_ak, lq_ak,
  by = "state"
)
dn_lq_ak
```

### Exercise 4

There are six observations in the joined dn_lq_ak data frame. The variables are the address, city, zip, longitude, and latitude of both Denny's and La Quinta and the state, which is Alaska.

```{r abt_merged_set}
dim(dn_lq_ak)
colnames(dn_lq_ak)
```

### Exercise 5

...

### Exercise 6

...

Add exercise headings as needed.
